Improving cropland maps through tight integration of human and machine intelligence

### 2019-10-04 manuscript/manuscript.Rmd
Updated manuscript and analyses (reference site selection)

Cropland mapping over a region such as Africa is difficult for several reasons. The primary reason relates to the characteristics of characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015]. This size is small relative to the 30-250 m resolution of the sensors typically used for most landcover mapping efforts, which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979], in this case the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons. These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.

@@

PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3). Imagery that don't having certain quality flags are excluded (e.g. failing to meet minimum spatial precision).

We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a reference grid that is situated within a larger 1 degree resolution grid that covers the entire continent. Each 1 degree cell defines the minimum mapping area of interest (Fig. 2A) is divided into 400 0.05 degree cells (Fig. 2B), which provides the scale for creating image composite tiles. The ARD process simply clips all available imagery intersecting each 0.05 cell to the extent of that cell, regardless of cloud cover.

@@

### Labelling platform
Both training and reference data are collected by a custom crowdsourcing platform, which refer to here for simplicity's sake as *mapper*. The original version of *mapper* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process. The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process. Each *mapper* instance is focused on a specific mapping geography (e.g. one to several 1 degree cells in Fig. 2A), thus several instances are set up to run in parallel to cover a larger mapping geography such as Ghana.

Here we provide an overview of *mapper*'s architecture within the context of the labeling workflow within a single instance.

#### Mapping workflow
##### Selecting training and reference sites
The labelling process begins with a python script that randomly selects a subset (e.g. 100) of the 0.005 degree grid (Fig. 2C) cells that intersect the *mapper* instance's specific area of interest (e.g. shaded grey cell in Fig. 2A). These site names are stored in a Postgres table that represents the mapping queue. Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.

@@

##### Processing completed assignments
Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table. The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning. If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score as based on the following formula:

@@

@@

Once the full set of training sites has been completed, two random draws of pre-determined size (typically **X%** each of the overall number of labels) are made from the randomly selected consensus labels. The first set is designated as a model test set, and the final set is set aside as final map validation sample. The remaining **Y%** of the labels are used to train the classifier.

### Classification pipeline
Upon completion of the labels, *mapper* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of memory-optimized instances (e.g. m4.xlarge), depending on the size of the mapping geography.

#### Feature extraction
The first step in the pipeline is the extraction of additional features from the growing season/non-growing season PlanetScope imagery. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work. Here, we used a set of X features, comprised of the eight original composite bands (four from each season), a 5X5 and YXY mean of each band, as well as the slope of each image calculated within the 5X5 and YXY windows (Table X).

@@

Feature         Window Size     N Features
-------------- ---------------- --------------------------------
RGB-NIR        1X1              8 
Low-pass       3X3              8
Slope          3X3              8
Other          YXY              X
-------------- ---------------- --------------------------------

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data. Features are extracted on the fly for each cell in the training and test sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.

#### Classification
Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of RandomForest [@BreimanRandomForests2001]. \textcolor{red}{The model is trained with a balanced sample and a tree depth of 16 and total tree number of 30, which initial testing showed to provide a reasonable balance between computational time and performance.}

After fitting, the model is applied to the features of the model test set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.

### The active learning loop
After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process. The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training, test, and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand. The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:

@@

After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to *mapper*, where workers then digitize new labels. These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning instance is launched and completed, and uncertainty scores for the remaining unlabelled sites are calculaetd, and the most uncertain sites are sent again to the workers for labeling. This loop repeats until the mapping skill as assessed with the model test set reaches an asymptote.

@@

To assess the accuracy of the final segmented boundaries, we used a two-step approach. First, we assessed the overall thematic accuracy of the resulting classification against our reserved validation set within each AOI. Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to map the given AOI. [**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because both an automated and  human-interpreted segmentation worker interpreting imagery cue in on--abrupt, physically detectable breaks within the imagery--are the same. On the other hand, no matter how well these boundaries align the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate such an approach against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).

@@

## The active learning process
We ran 16 instances to create cropland probability maps for the **X test regions/whole of Ghana**. Each instance covered a distinct mapping geography comprised of one or more adjacent AOIs (Figure 5), ranging in extent from 12,160 to 23,535 km$^2$ (average = 15,457 km$^2$), and represented a separate active learning process. The average number of iterations per active learning process was **X (range = X-X)**, and the average overall accuracy was **X, with a range of X-X** (Figure 5). Overall user's accuracy was slightly **lower/higher** at **X, (range X-X)**, while producer's accuracy averaged **X, (range X-X)**. Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southerwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.

>>>>

Cropland mapping over a region such as Africa is difficult for several reasons. The primary reason relates to the characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015]. This size is small relative to the 30-250 m resolution of the sensors typically used for most landcover mapping efforts, which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979], in this case the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons. These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.

@@

PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3).

We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent. Each 1 degree cell defines the minimum mapping area of interest (Fig. 2A) is divided into 400 0.05 degree cells (Fig. 2B), which provides the scale for creating image composite tiles. The ARD process simply clips all available imagery intersecting each 0.05 cell to the extent of that cell, regardless of cloud cover.

@@

### Labelling platform
Both training and reference data are collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*. The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process. The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process. Each *labeller* instance is focused on a specific mapping geography (e.g. one to several 1 degree cells in Fig. 2A), thus several instances are set up to run in parallel to cover a larger mapping geography such as Ghana.

Here we provide an overview of *labeller*'s architecture within the context of the labeling workflow within a single instance.

#### Mapping workflow
##### Selecting training and reference sites
The labelling process begins with a python script that randomly selects a subset (e.g. 100) of the 0.005 degree grid (Fig. 2C) cells that intersect the *labeller* instance's specific area of interest (e.g. shaded grey cell in Fig. 2A). These site names are stored in a Postgres table that represents the mapping queue. Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.

@@

##### Processing completed assignments
Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table. The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning. If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score as based on the following formula:

@@

Once a the full set of sites has been completed, they are randomly partitioned into training and validation sets according to predetermined proportions.

@@

### Classification pipeline
Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances (e.g. m4.xlarge), depending on the size of the mapping geography.

#### Feature extraction
The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work. Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table X).

@@

Feature             Window Size  N Features
------------------- ------------ ------------
RGB-NIR                      1X1            8 
Mean                       11X11            8
Standard deviation           5X5            8
------------------ ------------- ------------

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.

#### Classification
Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of RandomForest [@BreimanRandomForests2001]. \textcolor{red}{The model is trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.}

After fitting, the model is applied to the features of the model validation set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.

### The active learning loop
After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process. The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand. The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:

@@

After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to *mapper*, where workers then digitize new labels. These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning cluster is launched, resulting in a reassessment of uncertainty for unlabelled sites, a selection from the most uncertain sites, and a labelling of those sites by the workers. This loop repeats until the map accuracy metrics, as assessed against the validation set, show no further gains.

@@

To assess the accuracy of the final segmented boundaries, we used a two-step approach. First, we assessed the overall thematic accuracy of the resulting classification against our map reference data. Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map. [**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery. On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).

### Accuracy assessment
**[Put in somewhere above text about training reference samples]**
The accuracy assessment procedure described in section 2.2 is used strictly to assess performance gains in the active learning loop, according to suggested best practice guidelines [@ElmesAccountingtrainingdata2019]. To assess the accuracy of final map products, following these same guidelines we used an independent map reference sample. To develop this sample, we first made a random selection of grid cells within each mapping AOI, with the selected number scaled relative to the AOI. Next, a primary supervisor examined the image composites within each randomly selected cell and placed three rectangular polygons of ~0.1 ha area in areas that were fully contained within in one of three classes: cropland, non-cropland, potentially cropland or non-cropland. In cases where less than three classes were present within the cell, the supervisor placed polygons in proportion to the dominant class. For example, in the case where the entire cell was taken up by non-cropland, the supervisor placed all three polygons in non-cropland. If two-thirds of the cell was non-cropland, and one-third ambiguous, two polygons would be placed in non-cropland, and one in the ambiguous class. After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery. The four sets of labels were then combined and the overall agreement between them was calculated. This agreement provided the upper bound on knowable map accuracy.

## Application and evaluation of the system
[Description of map creation process]
To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each 1X1 degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent 1X1 cells that overlapped with neighboring countries. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.

Primary to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI). Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group. At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations resulted in the

To more fully evaluate the capacities of different components of the system, we performed several experiments. In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels. This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy, as recommended by @ElmesAccountingtrainingdata2019. Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites. Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.  Finally, we evaluated production performance in terms of speed of labelling and accuracy of results by comparing the generation time (label creation, training, and then classification) and accuracy of maps trained by 1) our core team of 10-12 experienced workers, and 2) a team of 100 novice workers from three different countries.

@@

## The active learning process
**[Fill in with intro]**. The average number of iterations per active learning process was **X (range = X-X)**, and the average overall accuracy was **X, with a range of X-X** (Figure 5). Overall user's accuracy was slightly **lower/higher** at **X, (range X-X)**, while producer's accuracy averaged **X, (range X-X)**. Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southerwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
